TensorAlloy: a highly efficient atomistic neural network program for alloys

### 2019-04-23 doc/papers/nn/manuscript.tex
Team revision is done. The version is CPC.v1.

\documentclass[prb,preprint]{revtex4-2}

@@

\title{
    TensorAlloy: a highly efficient atomistic neural network program for alloys
}
\author{Xin Chen, De-Ye Lin, Hai-Feng Song$^\mathrm{*}$}
\affiliation{
    Institute of Applied Physics and Computational Math, 
    Beijing 100088, China}

\affiliation{
    CAEP Software Center for High Performance Numerical Simulation, 
    Beijing 100088, China}

\begin{abstract}
Atomistic modeling is important for studying physical and chemical properties of
materials. Recently, machine learning interaction potentials have gained much 
more attentions as they can give density functional theory level predictions 
within second. Currently, symmetry function based atomistic neural network is 
the most widely used model for alloys. To precisely describe atomistic 
interactions, integrating advanced metrics, such as force or virial stress, into 
training can be of great help. However, the traditional way to construct a 
machine learning model computation graph starts from pre-computed descriptors, 
but not positions, due to technial challenges. Thus, deriving analytical atomic 
force or virial stress of a machine learning model becomes quite difficult as 
the partial derivations of descriptors with respect to positions need separate 
implementations, which requires plenty of effort and its efficiency can not be 
guaranteed.
In this paper, we designed a new method, named the virtual atom approach, for 
implementing the symmetry function descriptor and the corresponding atomistic 
neural networks. With the virtual atom approach, we are able to build 
computation graph from atomic positions \textemdash but not descriptors 
\textemdash to total energy directly, thus asking modern machine learning 
platforms to compute atomic force and virial stress becoming possible. We also 
derived a simple and machine learning friendly equation for calculating virial 
stress. All these new algorithms have been implemented in our Python package, 
TensorAlloy. TensorAlloy supports both \textit{molecules} and \textit{solids} 
and one hour is sufficient for TensorAlloy to obtain models with state-of-art 
level accuracy.
\end{abstract}

@@

Atomistic simulations (\textit{ab initio} calculation, classic molecular 
dynamics simulation (MD), etc) are powerful tools for studying physical and 
chemical properties of materials. To obtain reliable simulation results, 
atomistic interactions must be properly described. The very accurate 
\textit{ab initio} and quantum chemical methods have gain significant 
improvements over the past decades but their applicabilities are restricted by 
computation expenses. Physical model based empirical potentials, such as the 
embedded-atom method\cite{eam-1,eam-2,eam-3,eam-4,eam-5}, are much more popular 
for long-time simulation because of their reasonable accuraty and acceptable 
costs. However, designing empirical potentials and optimizing parameters for 
alloys are still challenging tasks.

In the past few years, machine learning (ML) has become one of the most 
influential topic in almost all fields of science. A lot of effort has been made 
by various physicists to explore the possiblity of using ML models \textemdash 
instead of traditional empirical potentials \textemdash to describe atomistic 
interactions during MD simulations. ML potentials are much easier to design and 
optimize. With sufficient training data, ML models can achieve similar or even 
better performances compared with density functional theory. Until now, various 
ML models (kCON\cite{kCON}, SchNet\cite{SchNet}, DTNN\cite{DTNN}, 
SNAP\cite{SNAP_Mo_2017, SNAP_2018}, etc) have been published by researchers. 
Among them, the symmetry function based atomistic neural network (ANN) model, 
first proposed by Parinello and Behler in 
2007\cite{Behler,Behler2,Behler3,Behler4,Behler1}, is still the most widely used 
method for \textit{solids} and its effectness and accuracy has been proven by 
several works\cite{Behler-solids-lzp-1,Behler-alloy-1,Behler-alloy-2,
Behler-alloy-3}. The Amp\cite{AMP} package, developed by Peterson's group, is 
the first open-source framework that implements the symmetry function 
descriptor. Amp supports both \textit{molecules} and \textit{solids} and it has 
successfully powered quite a few researches since 
2016\cite{Amp-works-hammer,amp-work-1}. ANI\cite{ANI} and 
TensorMol\cite{TensorMol} are also symmetry function based packages but they 
mainly focus on \textit{molecules}.

In order to precisely model atomistic interactions, both total energy and atomic 
forces are necessary. For \textit{solids}, the virial stress tensor is also an
essential metric. However, integrating force and stress into the loss function 
is not an easy task. The machine learning approaches favor vectorizable 
operations but the calculations of atomic descriptors are typically complicated 
so that many traditional ML packages (Amp, kCON, etc) choose to pre-compute 
descriptors and only build computation graph from descriptors to total energy. 
This approach can significantly reduce technical difficulty and works very well 
if only total energy is considered in the loss function. But to fit atomic 
forces, numerous codes must be written manually, such as the derivation of 
descriptors with respect to positions. Although the derivation is theoretically
clear, the implementation will be a challenge and the efficiency can not be 
guaranteed.

@@

This paper is organized as follows. Section \ref{section:theory} introduces the 
theory of the symmetry function descriptor and the ANN framework. Section 
\ref{section:method} describes the virtual atom approach and a modified ANN 
model used by \tensoralloy{}. Section \ref{section:discussions} displays 
performances of \tensoralloy{} on two public datasets. The appendix 
\ref{section:appendix} gives derivation and implementation details of key 
algorithms. The compiled version of \tensoralloy{} program and all the 
associated model files are available on 
\href{https://github.com/Bismarrck/tensoralloy-release}{GitHub}.

@@

In the atomistic neural network (\textbf{ANN}) framework, the total energy, 
$E^{total}$, of a structure with $N$ atoms, is the sum of all atomic energies:
\begin{equation}
\label{eq:general_e_total}
E^{total} = \sum_{i}^{N}{E_i}
\end{equation}
where $E_{i}$, the energy of atom \textit{i}, is the output of the neural 
network $\mathbf{NN}_{el}$ for element $el$:
\begin{equation}
E_i = \mathbf{NN}_{el}\left( \mathbf{G}_i \right)
\end{equation} 
Here $\mathbf{G}_i$ is a vector representing atomic descriptors of atom $i$. 
$\mathbf{G}_i$ typically only depends on local environments \textemdash 
neighbors \textemdash of atom $i$:
\begin{equation}
\mathbf{G}_i = F_G(\left\{ \rijn, |\rijn| < r_c\right\} )
\end{equation}
where $F_G$ can be arbitrary function, $r_c$ is the cutoff radius and $\rijn$ is 
the interatomic distance considering the periodic boundary conditions: 
\begin{equation}
    \label{eq:rijn}
    \rijn = \Vert \mathbf{r}_i^{(0)} - \mathbf{r}_j^{(0)} + 
    \mathbf{n}^T \mathbf{h} \Vert
\end{equation}
where $\mathbf{r}_i^{(0)}$ and $\mathbf{r}_j^{(0)}$ denote positions of $i$ and 
$j$ in the primitive cell, $\mathbf{n}$ is a column vector specifying the cell 
shifts along $(X,Y,Z)$ directions 
\begin{equation}
    \label{eq:ij_shift}
    \mathbf{n} = \begin{pmatrix}
        n_x \\
        n_y \\
        n_z
    \end{pmatrix}
\end{equation}
and $\mathbf{h}$ is the \textbf{row-major} $3 \times 3$ lattice tensor:
\begin{equation}
    \label{eq:lattice}
    \mathbf{h} = \begin{pmatrix}
        h_{xx} & h_{xy} & h_{xz} \\
        h_{yx} & h_{yy} & h_{yz} \\
        h_{zx} & h_{zy} & h_{zz} \\
    \end{pmatrix}
\end{equation}

@@

There are many choices of $F_G$. The symmetry function descriptor, proposed 
by Parinello and Behler in 2007\cite{Behler}, is currently the most widely used 
atomic descriptor, especially for alloys. The symmetry function descriptor 
consists of two sets of invariant (tranlational, rotational and permutational) 
functions: the radial symmetry function $\mathbf{G}^{(2)}$ and the angular 
symmetry function $\mathbf{G}^{(4)}$:
\begin{align}
\label{eq:G2}
\mathbf{G}^{(2)}_i(\eta, R_s) = & \sum_{j \neq i}{
    \sum_{\mathbf{n}}{g_2 (\eta, i, j, R_s)}
} \\
\label{eq:G4}
\mathbf{G}^{(4)}_i(\beta, \gamma, \zeta) 
= & 2^{1-\zeta}\sum_{j, k\neq j, k\neq i}{
    \sum_{\mathbf{n}_1}{\sum_{\mathbf{n}_2}{\sum_{\mathbf{n}_3}{
        g_4(\beta, \gamma, \zeta, 
          i, j, k, 
          \mathbf{n}_1, \mathbf{n}_2, \mathbf{n}_3)
    }}}
} \\
\label{eq:g2}
g_2 = & \exp\left\{-\frac{\eta(\rijn - R_s)^2}{r_c^2} \right\}
\cdot f_c(\rijn) \\
g_4 = & \left( 1 + \gamma\cos\theta \right)^\zeta 
    \exp \left\{ -\frac{\beta(\rijna^2 + \rikna^2 + \rjkna^2)}{r_c^2} \right\}
f_c(\rijna)f_c(\rikna)f_c(\rjkna)
\end{align}
where $\eta$, $R_s$, $\beta$, $\gamma$ and $\zeta$ are empirically chosen 
parameters, $f_c(r)$ is a cutoff (damping) function and its original form is:
\begin{equation}
\label{eq:fc_origin}
f_c(r) =
    \begin{cases}
    0 & \text{$r > r_c$} \\
    \frac{1}{2} + \frac{1}{2}\cos\left(r / r_c \cdot \pi \right) & 
    \text{$r \leq r_c$}
    \end{cases}
\end{equation}
and $r_c$ is the cutoff radius. In our implementation, $R_s$ is always fixed to 
0 so it becomes neglectable\cite{AMP}. Equation \eqref{eq:fc_origin} can be 
further transformed to a vectorized expression:
\begin{equation}
\label{eq:fc}
f_c(r) = \frac{1}{2}\left( 
    1 + \cos\left[ \min(\frac{r}{r_c}, 1) \pi \right] 
\right)
\end{equation} 
and $\cos\theta$ also has a vectorized form:
\begin{equation}
\cos\theta = \frac{\rijna^2 + \rikna^2 - \rjkna^2}{2 \rijna \cdot \rikna}
\end{equation}
The summations in $\mathbf{G}^{(2)}$ and $\mathbf{G}^{(4)}$ should go over all 
neighbors within $r_c$.

@@

The derivations are theoretically straightforward but coding these equations 
manually will be a challenge\cite{AMP}. Thanks to the virtual atom approach, 
\tensoralloy{} can now calculate $f_a$ using Equation \ref{eq:dEdr} directly.

The $3 \times 3$ virial stress tensor $\epsilon$ is an important metric for 
solids. For arbitrary many-body interaction potentials, $\epsilon$ can be 
calculated with the following equation\cite{lammps_stress}:
\begin{equation}
\label{eq:stress_lammps}
\epsilon = \left(-\sum_{i=1}^{N}{\mathbf{r}_i^{(0)} \otimes f_i} - 
\sum_{\mathbf{n}}{\mathbf{h}^T\mathbf{n}} \otimes 
\sum_{i=1}^{N}{F^{\prime}_{i\mathbf{n}}}
\right)^T
\end{equation}
where $\otimes$ indicates tensor product, $f_i$ is the total force acting on 
atom $i$ and $F^{\prime}_{i\mathbf{n}}$ is the partial force. Yntil now, 
most machine learning force-field models focus on fitting total energy and 
atomic forces, very few of them\cite{DeePMD,DeePMD_rl,DeePMD_kit} had attempted 
to include stress in their loss functions. The current expression of Equation 
\ref{eq:stress_lammps} is not very compatible with NN or other ML potentials 
since the partial forces are difficult to compute with vectorized operations. 
To resolve this issue, we derived an equivalent form of Equation 
\ref{eq:stress_lammps}:
\begin{equation}
\label{eq:stress}
\epsilon = -F^{T} R + 
\left(\frac{\partial E^{total}}{\partial \mathbf{h}}\right)^T \mathbf{h}
\end{equation}
where $R$ is the $N \times 3$ positions matrix and $F$ is the corresponding 
total forces matrix. The new equation can be easily implemented with simple 
matrix operations. The detailed derivation is given in the appendix.

@@

In this section we will describe details of \tensoralloy{}: the virtual atom 
approach, the atomic residual potential and the loss function. The Ni-Mo alloy 
system, with four $\eta$, two $\beta$, two $\gamma$ and one $\zeta$, will be 
used as the example to show this algorithm. For simplicity, only radial symmetry
function related algorithms are visualized in the figures.

@@

Fig \ref{fig:algo1} shows the initialization procedure of the virtual atom 
approach. The cutoff radius $r_{cut}$ must be fixed. The first step is to go 
through the entire dataset and determining four prerequisite constants: 
$N_{ij}^{\mathrm{max}}$, $N_{ijk}^{\mathrm{max}}$, 
$N_{\mathrm{Ni}}^{\mathrm{max}}$ and $N_{\mathrm{Mo}}^{\mathrm{max}}$. 
Here $N_{ij}^{\mathrm{max}}$ represents the maximum number of neighbor pairs of 
arbitrary structure and $N_{ijk}^{\mathrm{max}}$ represents the maximum number 
of triples. $N_{ij}^{\mathrm{max}}$ and $N_{ijk}^{\mathrm{max}}$ only depends on 
$r_{cut}$. The associated codes are given in the appendix.
$N_{\mathrm{Ni}}^{\mathrm{max}}$ and $N_{\mathrm{Mo}}^{\mathrm{max}}$ indicate 
the maximum appearances of Ni and Mo in any structure, respectively. 
With $N_{\mathrm{Ni}}^{\mathrm{max}}$ and $N_{\mathrm{Mo}}^{\mathrm{max}}$, 
the global symbol list (GSL) can be created. GSL defines the global reference 
of this dataset and must be maintained during entire training. Then we can 
construct \textit{offsets}. This vector marks the starting indices of the 
\textbf{ordered} interaction pairs (e.g. Ni-Mo) and triples (e.g. Ni-Ni-Mo). 
The order of the interactions must be maintained as well.

\begin{figure}[h!]
\centering
\includegraphics[scale=0.5]{figures/Fig1-prepare.pdf}
\caption{\label{fig:algo1} The initialization. 
\textbf{a)} the detections of $N_{\mathrm{Ni}}^{\mathrm{max}}$, 
$N_{\mathrm{Ni}}^{\mathrm{max}}$, $N_{ij}^{max}$, $N_{ijk}^{max}$ and the global
symbols list.
\textbf{b)} and \textbf{c)} the \textit{offsets} and their corresponding 
symmetry function parameters.
}
\end{figure}

@@

The following equation briefly demonstrates the core data flow of 
\textbf{TensorAlloy} during a mini-batch training: 
\begin{equation}
    \mathbf{R'_{b}} \rightarrow 
    \mathbf{G_b} \rightarrow 
    \mathbf{E_b} \rightarrow
    \mathbf{F_b} \rightarrow 
    \mathbf{\epsilon_b} \rightarrow
    \mathbf{Loss}
\end{equation}
where the $N_{b} \times (\nmax + 1) \times 3$ matrix $\mathbf{R'_b}$ is a batch 
of positions matrices ($N_b \ge 1$ is the batch size), $\mathbf{G_b}$ is the 
corresponding symmetry function descriptors matrix of shape 
$N_{b} \times (\nmax) \times 16$ and $\mathbf{E_b}$ represents 
total energies of the structures. Section \ref{section:virtual_atom_approach} 
describes the implementation of $\mathbf{R'_{b}} \rightarrow \mathbf{G_b}$. 
To implement $\mathbf{G_b} \rightarrow \mathbf{E_b}$, \textbf{TensorAlloy} uses 
1D-convolutional neural networks with $1 \times 1$ kernels\cite{kCON}. The total 
forces $\mathbf{F_{b}}$ and the virial stress $\mathbf{\epsilon_{b}}$ are 
calculated with Equation \ref{eq:dEdr} and Equation \ref{eq:stress} by with just 
few lines of codes.

@@

To fit the total energy, we slightly modified the total energy expression of 
Equation \ref{eq:general_e_total}:
\begin{align}
\label{eq:e_total_res}
E^{total} & = E^{residual} + E^{static} \nonumber \\
& = \sum_{i=1}^{N}{
    \mathbf{NN}_{el}\left( \left\{ G_i \right\} \right)
} + \sum_{el}{n_{el}E_{el}}
\end{align}
where $\{E_{el}\}$ is a set of \textit{trainable} scalar variables representing 
the \textit{static} energy of each type of element\cite{kCON}. The NN in 
Equation \ref{eq:e_total_res} just describes the atomistic interactions 
(the \textit{residual} energy). The 
introduction of $E^{static}$ can significantly limit value range of NN outputs, 
leading to much faster convergence speed and higher training stability 
(Fig \ref{fig:benchmark_qm7}c). The initial $\{E_{el}\}$ are calculated by 
solving the linear system $Ax=b$ where $A$ is a $N_{data} \times N_{el}$ matrix 
and $b$ is a $N_{data}\times 1$ vector. $A(i,j)$ is the number of $j^{th}$
element in structure $i$ and $b(i)$ is the total energy of structure $i$.

>>>>

\documentclass[preprint]{revtex4-2}

@@

\title{
    TensorAlloy: a highly efficient atomistic neural network program for alloys
}
\author{Xin Chen}
\author{Xing-Yu Gao}
\author{Ya-Fan Zhao}
\author{De-Ye Lin}
\email{lindeye0716@163.com}
\author{Wei-Dong Chu}
\author{Hai-Feng Song}
\email{song\_haifeng@iapcm.ac.cn}
\affiliation{
    Institute of Applied Physics and Computational Math, 
    Beijing 100088, China}
\affiliation{
    CAEP Software Center for High Performance Numerical Simulation, 
    Beijing 100088, China}

\begin{abstract}
Atomistic modeling is important for studying physical and chemical properties of
materials. Recently, machine learning interaction potentials have gained much 
more attentions as they can provide density functional theory level predictions 
within negligible time. Currently, the symmetry function descriptor based 
atomistic neural network is the most widely used model for modeling alloys. 
To precisely describe complex potential energy surfaces, integrating advanced 
metrics, such as force or virial stress, into training can be of great help. 
However, the traditional way to construct a 
machine learning model computation graph starts from pre-computed descriptors, 
but not positions, due to technial challenges. Thus, deriving analytical atomic 
force and virial stress of a machine learning model becomes quite difficult as 
the partial derivations of descriptors with respect to positions need separate 
implementations, which requires plenty of efforts and its efficiency can not be 
guaranteed.
In this paper, we designed a new method, named the virtual atom approach, for 
implementing the symmetry function descriptor and the corresponding atomistic 
neural networks. With the virtual atom approach, we are able to build 
computation graph from atomic positions \textemdash but not descriptors 
\textemdash to total energy directly, thus asking modern machine learning 
platforms to compute atomic force and virial stress becoming possible. We also 
derived a simple and machine learning friendly equation for calculating virial 
stress. All these new algorithms have been implemented in our Python package, 
TensorAlloy. TensorAlloy supports constructing machine learning interaction 
potentials for both \textit{molecules} and \textit{solids} and one hour is 
sufficient for TensorAlloy to obtain models with state-of-art level accuracy.
\end{abstract}

@@

Atomistic simulations (\textit{ab initio} calculation, classic molecular 
dynamics simulation (MD), etc) are powerful tools for studying physical and 
chemical properties of materials. To obtain reliable simulation results, 
atomistic interactions must be properly described. The very accurate 
\textit{ab initio} and quantum chemical methods have gain significant 
improvements over the past decades but their applicabilities are restricted by 
computation expenses. Physical model based empirical potentials, such as the 
embedded-atom method\cite{eam-1,eam-2,eam-3,eam-4,eam-5} or its variants (
modified embedded-atom method\cite{eam-5,meam-1,meam-2,meam-3}, 
angular-dependent interaction potential\cite{adp-1,adp-2,adp-3}, etc), are much 
more popular for long-time simulation because of their reasonable accuraty and 
acceptable costs. 
However, designing empirical potentials and optimizing parameters for 
alloys are still challenging tasks\cite{eam-5, adp-2, SNAP_2018}.

In the past few years, machine learning (ML) has become one of the most 
influential topic in almost all fields of science. A lot of effort has been made 
by various physicists to explore the possiblity of using ML models \textemdash 
instead of traditional empirical potentials \textemdash to describe atomistic 
interactions during MD simulations\cite{kCON,SchNet,DTNN,SNAP_2018,SNAP_Mo_2017,
ANI,TensorMol}. ML potentials are much easier to design and 
optimize. With sufficient training data, ML models can achieve similar or even 
better performances compared with density functional theory.
The symmetry function based atomistic neural network (ANN) model, first proposed 
by Parinello and Behler in 2007\cite{Behler,Behler2,Behler3,Behler4,Behler1}, 
is still the most widely used method for \textit{solids} and its effectness and 
accuracy has been proven by several 
works\cite{Behler-solids-lzp-1,Behler-alloy-1,Behler-alloy-2, Behler-alloy-3}.
Until now, various ML implementations\cite{AMP,ANI,TensorMol} have been 
published by researchers. The Amp\cite{AMP} package, developed by Peterson's 
group, is the first open-source framework that implements the symmetry function 
descriptor. Amp can be used to learn interaction potentials for both 
\textit{molecules} and \textit{solids} and it has successfully powered quite a 
few researches since 2016\cite{Amp-works-hammer,amp-work-1}. ANI\cite{ANI} and 
TensorMol\cite{TensorMol} are also symmetry function based packages but they 
mainly focus on \textit{molecules}.

In order to precisely model atomistic interactions, both total energy and atomic 
forces are necessary. For \textit{solids}, the virial stress tensor is also an
essential metric. However, integrating force and stress into the loss function 
is not an easy task. The machine learning approaches favor vectorizable 
operations but the calculations of atomic descriptors are typically complicated 
so that many traditional ML packages (Amp, kCON, etc) choose to pre-compute 
descriptors and only build computation graph from descriptors to total energy. 
This approach can significantly reduce technical difficulty and works very well 
if only total energy is considered in the loss function. But to fit atomic 
forces, numerous codes must be written manually, such as the derivation of 
descriptors with respect to positions. Although the derivation is theoretically
clear, the implementation is challenging and the efficiency can not be 
guaranteed.

@@

This paper is organized as follows. Section \ref{section:theory} introduces the 
theory of the symmetry function descriptor and the ANN framework. Section 
\ref{section:method} describes the virtual atom approach and a modified ANN 
model used by \tensoralloy{}. Section \ref{section:discussions} displays 
performances of \tensoralloy{} on two public datasets. The appendix 
\ref{section:appendix} gives derivation and implementation details of key 
algorithms. The \tensoralloy{} program can be obtained on request.

@@

In the atomistic neural network (\textbf{ANN}) framework, the total energy, 
$E^{total}$, of a structure with $N$ atoms, is the sum of all atomic energies:
\begin{equation}
\label{eq:general_e_total}
E^{total} = \sum_{i}^{N}{E_i}
\end{equation}
where $E_{i}$, the energy of atom \textit{i}, is the output of the neural 
network $\mathbf{NN}_{el}$ for element $el$:
\begin{equation}
E_i = \mathbf{NN}_{el}\left( \mathbf{G}_i \right)
\end{equation} 
Here $\mathbf{G}_i$ is a vector representing atomic descriptors of atom $i$. 
$\mathbf{G}_i$ typically only depends on local environments, i.e.  
neighbors of atom $i$:
\begin{equation}
\mathbf{G}_i = F_G(\left\{ \rijn, |\rijn| < r_c\right\} )
\end{equation}
where $F_G$ can be arbitrary function, $r_c$ is the cutoff radius and $\rijn$ is 
the interatomic distance considering the periodic boundary conditions: 
\begin{equation}
    \label{eq:rijn}
    \rijn = \Vert \mathbf{r}_i^{(0)} - \mathbf{r}_j^{(0)} + 
    \mathbf{n}^T \mathbf{h} \Vert
\end{equation}
where $\mathbf{r}_i^{(0)}$ and $\mathbf{r}_j^{(0)}$ denote positions of $i$ and 
$j$ in the primitive cell, $\mathbf{n}$ is a column vector specifying the cell 
shifts along $(X,Y,Z)$ directions 
\begin{equation}
    \label{eq:ij_shift}
    \mathbf{n} = \begin{pmatrix}
        n_x \\
        n_y \\
        n_z
    \end{pmatrix}
\end{equation}
and $\mathbf{h}$ is the \textbf{row-major} $3 \times 3$ lattice tensor:
\begin{equation}
    \label{eq:lattice}
    \mathbf{h} = \begin{pmatrix}
        h_{xx} & h_{xy} & h_{xz} \\
        h_{yx} & h_{yy} & h_{yz} \\
        h_{zx} & h_{zy} & h_{zz} \\
    \end{pmatrix}
\end{equation}

@@

There are many choices of $F_G$. The symmetry function descriptor, proposed 
by Parinello and Behler in 2007\cite{Behler}, is currently the most widely used 
atomic descriptor, especially for alloys. The symmetry function descriptor 
consists of two sets of invariant (tranlational, rotational and permutational) 
functions: the radial symmetry function $\mathbf{G}^{(2)}$ and the angular 
symmetry function $\mathbf{G}^{(4)}$:
\begin{align}
\label{eq:G2}
\mathbf{G}^{(2)}_i(\eta, R_s) = & \sum_{j \neq i}{
    \sum_{\mathbf{n}}{g_2 (\eta, i, j, R_s)}
} \\
\label{eq:G4}
\mathbf{G}^{(4)}_i(\beta, \gamma, \zeta) 
= & 2^{1-\zeta}\sum_{j, k\neq j, k\neq i}{
    \sum_{\mathbf{n}_1}{\sum_{\mathbf{n}_2}{\sum_{\mathbf{n}_3}{
        g_4(\beta, \gamma, \zeta, 
          i, j, k, 
          \mathbf{n}_1, \mathbf{n}_2, \mathbf{n}_3)
    }}}
} \\
\label{eq:g2}
g_2 = & \exp\left\{-\frac{\eta(\rijn - R_s)^2}{r_c^2} \right\}
\cdot f_c(\rijn) \\
g_4 = & \left( 1 + \gamma\cos\theta \right)^\zeta 
    \exp \left\{ -\frac{\beta(\rijna^2 + \rikna^2 + \rjkna^2)}{r_c^2} \right\}
f_c(\rijna)f_c(\rikna)f_c(\rjkna)
\end{align}
where $\eta$, $R_s$, $\beta$, $\gamma$ and $\zeta$ are empirically chosen 
parameters, $f_c(r)$ is a cutoff (damping) function and its original form is:
\begin{equation}
\label{eq:fc_origin}
f_c(r) =
    \begin{cases}
    0 & \text{$r > r_c$} \\
    \frac{1}{2} + \frac{1}{2}\cos\left(r / r_c \cdot \pi \right) & 
    \text{$r \leq r_c$}
    \end{cases}
\end{equation}
and $r_c$ is the cutoff radius. $R_s$ may be fixed to zero \cite{AMP}. 
Equation \eqref{eq:fc_origin} can be further transformed to a vectorized 
expression:
\begin{equation}
\label{eq:fc}
f_c(r) = \frac{1}{2}\left( 
    1 + \cos\left[ \min(\frac{r}{r_c}, 1) \pi \right] 
\right)
\end{equation} 
and $\cos\theta$ also has a vectorized form:
\begin{equation}
\cos\theta = \frac{\rijna^2 + \rikna^2 - \rjkna^2}{2 \rijna \cdot \rikna}
\end{equation}
The summations in $\mathbf{G}^{(2)}$ and $\mathbf{G}^{(4)}$ should go over all 
neighbors within $r_c$.

@@

The derivations are theoretically straightforward but their implementations are 
challenging \cite{AMP}. Later we will provide a new approach to address this 
problem.

The $3 \times 3$ virial stress tensor $\epsilon$ is an important metric to 
describe solids under deformation or external pressure. For arbitrary many-body 
interaction potentials, $\epsilon$ can be calculated with the following equation
\cite{lammps_stress}:
\begin{equation}
\label{eq:stress_lammps}
V \cdot \epsilon = \left(-\sum_{i=1}^{N}{\mathbf{r}_i^{(0)} \otimes f_i} - 
\sum_{\mathbf{n}}{\mathbf{h}^T\mathbf{n}} \otimes 
\sum_{i=1}^{N}{F^{\prime}_{i\mathbf{n}}}
\right)^T
\end{equation}
where $V$ is the volume, $\otimes$ indicates tensor product, $f_i$ is the total 
force acting on atom $i$ and $F^{\prime}_{i\mathbf{n}}$ is the partial force. 
Until now, most machine learning force-field models focus on fitting total 
energy and atomic forces, very few of them\cite{DeePMD,DeePMD_rl,DeePMD_kit} had 
attempted to include stress in their loss functions. 
The current expression of Equation \ref{eq:stress_lammps} is not very compatible 
with NN or other ML potentials since the partial forces are difficult to compute 
with vectorized operations. To resolve this issue, we derived an equivalent form 
of Equation \ref{eq:stress_lammps}:
\begin{equation}
\label{eq:stress}
V \cdot \epsilon = -F^{T} R + 
\left(\frac{\partial E^{total}}{\partial \mathbf{h}}\right)^T \mathbf{h}
\end{equation}
where $R$ is the $N \times 3$ positions matrix and $F$ is the 
corresponding total forces matrix. The new equation can be easily implemented 
with simple matrix operations. The detailed derivation is given in the appendix.

@@

In this section we will describe details of \tensoralloy{}: the virtual atom 
approach, the atomic residual potential and the loss function. 
The Ni-Mo binary alloy system will be used as the example to show this 
algorithm. The virtual atom approach is built upon TensorFlow. For simplicity, 
only radial symmetry function related algorithms are visualized in the figures.

@@

Fig \ref{fig:algo1} shows the initialization procedure of the virtual atom 
approach. The cutoff radius $r_{cut}$ must be fixed. The first step is to go 
through the entire dataset and determining four prerequisite constants: 
$N_{ij}^{\mathrm{max}}$, $N_{ijk}^{\mathrm{max}}$, 
$N_{\mathrm{Ni}}^{\mathrm{max}}$ and $N_{\mathrm{Mo}}^{\mathrm{max}}$. 
Here $N_{ij}^{\mathrm{max}}$ represents the maximum number of neighbor pairs 
and $N_{ijk}^{\mathrm{max}}$ represents the maximum number of triples in 
arbitrary structure. 
$N_{ij}^{\mathrm{max}}$ and $N_{ijk}^{\mathrm{max}}$ only depend on 
$r_{cut}$. The associated codes are given in the appendix.
$N_{\mathrm{Ni}}^{\mathrm{max}}$ and $N_{\mathrm{Mo}}^{\mathrm{max}}$ indicate 
the maximum appearances of Ni and Mo in any structure, respectively. 
With $N_{\mathrm{Ni}}^{\mathrm{max}}$ and $N_{\mathrm{Mo}}^{\mathrm{max}}$, 
the global symbol list (GSL) can be created. GSL defines the global reference 
of this dataset and must be maintained during entire training. Then we can 
construct \textit{offsets}. This vector marks the starting indices of the 
\textbf{ordered} interaction pairs (e.g. Ni-Mo) and triples (e.g. Ni-Ni-Mo). 
The order of the interactions must be maintained as well.

\begin{figure}[h!]
\centering
\includegraphics[scale=0.5]{figures/Fig1-prepare.pdf}
\caption{\label{fig:algo1} The initialization step. 
\textbf{a)} describes the detections of $N_{\mathrm{Ni}}^{\mathrm{max}}$, 
$N_{\mathrm{Ni}}^{\mathrm{max}}$, $N_{ij}^{max}$, $N_{ijk}^{max}$ and the global
symbols list.
The \textit{offsets} of radial and angular functions and their corresponding 
symmetry function parameters are in \textbf{b)} and \textbf{c)}, respectively.
}
\end{figure}

@@

The following equation briefly demonstrates the core data flow of 
\textbf{TensorAlloy} during a mini-batch training: 
\begin{equation}
    \mathbf{R'_{b}} \rightarrow 
    \mathbf{G_b} \rightarrow 
    \mathbf{E_b} \rightarrow
    \mathbf{F_b} \rightarrow 
    \mathbf{\epsilon_b} \rightarrow
    \mathbf{Loss}
\end{equation}
where the $N_{b} \times (\nmax + 1) \times 3$ matrix $\mathbf{R'_b}$ is a batch 
of positions matrices ($N_b \ge 1$ is the batch size), $\mathbf{G_b}$ is the 
corresponding symmetry function descriptors matrix of shape 
$N_{b} \times (\nmax) \times 16$ and $\mathbf{E_b}$ represents 
total energies of the structures. Section \ref{section:virtual_atom_approach} 
describes the implementation of $\mathbf{R'_{b}} \rightarrow \mathbf{G_b}$. 
To implement $\mathbf{G_b} \rightarrow \mathbf{E_b}$, \textbf{TensorAlloy} uses 
1D-convolutional neural networks with $1 \times 1$ kernels\cite{kCON}. The total 
forces $\mathbf{F_{b}}$ and the virial stress $\mathbf{\epsilon_{b}}$ are 
calculated with Equation \ref{eq:dEdr} and Equation \ref{eq:stress} directly.

@@

To fit the total energy, we slightly modified the total energy expression of 
Equation \ref{eq:general_e_total}:
\begin{align}
\label{eq:e_total_res}
E^{total} & = E^{residual} + E^{static} \nonumber \\
& = \sum_{i=1}^{N}{
    \mathbf{NN}_{el}\left( \left\{ G_i \right\} \right)
} + \sum_{el}{n_{el}E_{el}}
\end{align}
where $\{E_{el}\}$ is a set of \textit{trainable} scalar variables representing 
the \textit{static} energy of each type of element\cite{kCON}. The NN in 
Equation \ref{eq:e_total_res} just describes the atomistic interactions 
(the \textit{residual} energy). The 
introduction of $E^{static}$ can significantly limit value range of NN outputs, 
leading to faster convergence speed and higher training stability 
(Fig \ref{fig:benchmark_qm7}c). The initial $\{E_{el}\}$ are calculated by 
solving the linear system $Ax=b$ where $A$ is a $N_{data} \times N_{el}$ matrix 
and $b$ is a $N_{data}\times 1$ vector. $A(i,j)$ is the number of $j^{th}$
element in structure $i$ and $b(i)$ is the total energy of structure $i$.
